\documentclass{article}
\usepackage{graphicx} % Required for inserting images
%\usepackage{geometry}
%\geometry{margin=0.5in}
%\usepackage{cite}
%\usepackage{color}

%\title{XMLTDB01}
%\author{Bo Sundman et al. }
%\date{June 2023}

\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{graphicx,subfigure}              % with figures
% sometimes needed to have pdf files 
\pdfsuppresswarningpagegroup=1
\topmargin -10mm
\oddsidemargin -10mm
\evensidemargin -10mm
\textwidth 170mm
\textheight 220mm
\parskip 2mm
\parindent 3mm
%\pagestyle{empty}
\usepackage{xcolor}
\usepackage[normalem]{ulem}

%footnote using capital letter
\renewcommand{\thefootnote}{\Alph{footnote}}
%\renewcommand{\thefootnote}{{\arabic{footnote}}

% For appendices
\usepackage[titletoc,title,header]{appendix}

\begin{document}

\begin{center}

  {\Large \bf A proposal of an XML format for TDB files}

  Bo Sundman et al. \today
\end{center}

\bigskip
{\bf Background}

The current TDB format for Calphad databases has worked well for 40
years but introducing new models and features has revealed some
awkward features, in particular the TYPE\_DEFINITION keyword.  When a
change is needed it may be interesting to adopt an accepted markup
language as XML or JSON and as it is easy to convert between any
generally accepted markup language, the XML has been selected as it is
slightly more explicit for a new XTDB format.

As XML is extendable each software may add their own flavour in a
controlled way and it will be simpler to use an XTDB file by different
softwares as one can indicate what is software specific.

The proposal here is a minimal one intended to:
\begin{itemize}
  \item Make manually editing an XTDB file as easy as the current TDB
    format.
  \item Minimize problems when using software A to read a database
    file written by software B.
  \item Provide better facilities to handle new models and features
    added by the different softwares using Calphad databases.
    Significant new features should be discussed with SGTE before
    integrated in the XTDB format.
\end{itemize}

This XTDB definition may require minor modifications in the XML format
already adoped in some software.  Commercial companies providing
encrypted databases to customers can use whatever format they prefer
for such databases.  But they should allow users to read and write
unencrypted databases in the XTDB format.


\bigskip
{\bf How to add comments to this proposal}

As we are using different editing facilities I have generated this PDF
with limited amount of text on each page and at the end of each page
in the attached LaTeX file (.tex) you are invited to add comments and
send it back to me.  Add a summary of your comments on the last page.

\% Add any comments below

\%---------------------------------------------------------------
%\newpage (without the \%)

Please add your comments using your preferred text editor without
changing any of the LaTeX formatting.

You are encoraged to ask anyone you think can contribute to defining
the new database format to have a look at this and provide their
comments.

In the table at the last page you can summarize your comments.

% Add any comments below

%---------------------------------------------------------------
\newpage

\section{XTDB version 0.0.1}

The following proposal is for a common definition of an XTDB file
format based on XML.  It can hopefully become an SGTE standard
acceptable by all software which develop or use Calphad databases.

The ``note'' for the attributes for an XML element in the tables below
is ``+'' for a compulsory attribute, ``--'' for an optional one or a
letter A-Z for special cases explained in section~\ref{sec:notes}.

All {\bf Id} attributes are case insenitive as in the TDB file.
Except for phase names the value of an {\bf Id} cannot be abbreviated
when used in other attributes.

An XML parser is convenient for reading XML files but to simplify the
life for the humans reading the XTDB file or software using sequential
read operations, it is recommended that the attributes to an XML
element are provided in the order listed for each element below.
Software specific attributes will come at the end.

The order of XML elements is free but to simplify for a human reading
and editing the XTDB file a recommended order is {\bf Defaults,
  DatabaseInfo, SoftSpecific} followed by {\bf Element, Species,
  Model} and {\bf Phases} and then {\bf TPfun}s and {\bf Parameter}s
for the pure elements.  After that {\bf TPfun} and {\bf Parameter} for
binary, ternary and higher order systems, normally arranged for each
subsystem.

An XML element referenced in the text below is written in {\bf bold}.

\subsection{The XTDB file}

The XTDB file should be divided into lines (terminated by LF (Line
Feed or newline as preferred by UNIX dialects) or CRLF (Carriage Return
and Line Feed as preferred by Windows)).  A line should not exceed
2000 characters.

Only one XML element per line (to be easily readable by humans) but an
XML element may extend over several lines, including nested elements.

Maybe all attributes of an XML element should fit within a line (2000
characters)?  Only nested XML elements on following lines.

For an example of a very long attribute see~\ref{sec:phase}.

A software reading the XTBD file should ignore XML elements and
attributes it does not support, preferably with a warning to the user.

% Add any comments below


%---------------------------------------------------------------
\newpage

\subsection{The system XTDB elements}

\bigskip
\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

  XTDB    & && Containing XML elements for an XTDB database.\\
          & Version & + & Version of XTDB for this file.\\
          &Software & + & Name of software generating the database.\\
          &Date     & + & Year/month/day the database was written or edited\\
          &Signature& -- & Name/email of person or organisation generating the database.\\\hline
  
  Defaults & && Provides default values of attributes in other XML elements.\\
           & LowT & -- & Default value of low $T$ limit.\\
           & HighT & -- & Default value of high $T$ limit.\\
           & Refid & -- & Default bibliographic reference. \\
           & Elements & -- & VA and/or /- (the electron).\\
           & Model & -- & Any model applicable for the system, for example EEC\\\hline

  DatabaseInfo & && Free text.\\
  & Date & -- & Last update.\\\hline

\end{tabular}

% Add any comments below

%---------------------------------------------------------------
\newpage

\subsection{The chemical element and species element}

\bigskip
\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

  Element & && Specifies a chemical element in the database.  The vacancy is denoted ``VA'' and the electron ``/-''.\\
          & Id        & +    & Chemical element symbol, for example FE, H \\
          & Refstate  & +    & Name of a phase, for example GAS \\
          & Mass      & +    & Mass in g/mol\\
          & H298      & --    & Enthalpy difference between 0 and 298.15~K in reference state.\\
          & S298      & --    & Entropy difference between 0 and 298.15~K in reference state.\\\hline

  Species & && Specifies a molecular like aggregate used as constituent of phases.  The elements, except the electron, are also species.\\
          & Id        & A & Species name max 24 letters. For note see \ref{sec:notes}.\\
          & Stoichiometry & B & One or more elements each followed by an unsigned real representing the stoichiometric ratio.  For note see \ref{sec:notes}.  See also appendix~\ref{sec:elementexample}.\\
          & MQMQA & C & For a constituent in the MQMQA model.  For note see \ref{sec:notes}.\\
          & UNIQUAC & D & For a constituent in the UNIQUAC model.  For note see \ref{sec:notes}.\\\hline

\end{tabular}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{The function and temperature range elements}

There is at present no way to handle several pressure ranges.  The
value of a function, as well as its first and second derivative with
respect to $T$, must be continuous across an interval of $T$ range.

\bigskip
\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

  TPfun & && Defines a $T, P$ expression to be used in parameters or other functions.\\
        & Id & E & Function name, max 16 characters. The name can be used in the ``Expr'' attribute of other functions or parameters.  For note see \ref{sec:notes}.\\
        & LowT & -- & If the default low $T$ limit applies.\\
        & Expr & + & Simple mathematical expression terminated by ;.   Use the {\bf Trange} element if several ranges.  See section \ref{sec:expr}\\
        & HighT & -- & If the default high $T$ limit applies.\\\hline

  Trange & && Only inside a {\bf TPfun} or {\bf Parameter} element for an expression with several $T$ ranges.\\ 
         & Expr & + & Simple mathematical expression terminated by ;.  See section \ref{sec:expr}.\\
         & HighT & -- & If the default high $T$ limit applies.\\\hline
\end{tabular}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{The phase element}\label{sec:phase}

Normally the software requires information of sublattices,
constituents and models for the phase in order to create the
appropriate data structures.  The {\bf Parameter} and some models, see
section~\ref{sec:toop} can appear anywhere in the XTDB file.

The {\bf Splitphase} model (also known as the disordered fraction
model) combines the Gibbs energies of an ordered and a disordered
phase where the disordered phase has fewer sublattices than the
ordered and the constituent fractions on its first sublattice are the
average of the fractions summed over several sublattices in the
ordered phase as specified by the attribute {\bf Sum}.  The {\bf
  SplitPhase} element must appear within the ordered {\bf phase}
element before any other nested XML elements such as {\bf AmendPhase,
  Parameter}, {\bf KohlerModel} or {\bf ToopModel} elements.  Tho {\bf
  Parameter, ToopModel} and {\bf KohlerModel} elements may appear anywhere in
the XTB file.

Some software may not have a separate disordered phase but arrange all
{\bf Parameter} elements for the ordered and disordered phase within
the same phase.

A phase with very long list of constituents (i.e. GAS) can have
several {\bf Constituent} elements for the same sublattice.

\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

  Phase & && All data belongs to a phase.\\
       & Id & F & Phase name, see \ref{sec:notes} note F. \\
       & Configuration & G & Model for the configurational entropy.  For note see \ref{sec:notes}.\\
       & State & -- & G for gas phase, L for liquid phase.  Needed if EEC is used.\\\hline

  Sublattices & && Only inside a {\bf Phase} element.\\
        & NumberOf & + & Number of sublattices, an integer value 1-10 \\
        & Multiplicities & + & Sites on each sublattice, as many reals as sublattices separated by a space.\\\hline

  Constituents & && Only inside a {\bf Phase} element.\\
        & Sublattice & + & Indicates the sublattice for the list of species.\\
        & List & + & Species Id, separated by a space, in the sublattice.\\\hline

  SplitPhase & && Only inside the {\bf Phase} element of the ordered phase.\\
      & Refid & + & Where the model is described.\\
%      & Ordered & J & May be omitted, for note see \ref{sec:notes}.\\
      & Disordered & K & Name of disordered phase, for note see \ref{sec:notes}.\\
      & Sum & + & Number of sublattices in the ordered phase to be summed for the disordered phase constitution.  All sublattices summed must have the some constituents.\\
      & Subtract & L & Must have value Y if the Gibbs energy of ordered phase as disordered should be subtracted, for note see \ref{sec:notes}.\\\hline

  AmendPhase & && Only inside a {\bf Phase} element.\\
  & Model & + & List of models ``Id'', separated by a space, for this phase.  The ``Id'' is specified in a model element below.\\\hline
\end{tabular}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{The simple parameter element}

All thermodynamic data are defined by the parameter elements.  They
can be arranged inside a phase element or separately for each binary,
ternary etc. subsystem.  See also section~\ref{sec:parameter2} and
  appendix~\ref{sec:parameter examples}.

\bigskip
\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

  Parameter & && Specifies the $T, P$ expression of a model parameter for a set of constituents.\\
      & Id & + & As in TDB file, for example G(LIQUID,A,B:VA;2).  See also the {\bf Parameter2} element.  May include the {\bf Trange} element for parameters with multiple ranges.\\
      & LowT & -- & If the default low $T$ limit applies.\\
      & Expr & + & Simple mathematical expression terminated by ;.  If several ranges use a {\bf Trange} element.  See section \ref{sec:expr}.\\
      & HighT & -- & If the default high $T$ limit applies.\\
      & Bibref & + & Bibliographic reference.\\\hline
\end{tabular}

The {\bf Id} starts with a model parameter identifier (MPID) followed
within parenthesis by a phase name and one or more constituents in
each sublattice of the phase.  Constituents in the same sublattice
(for interaction parameters) are separated by a comma, ``,'', and a
colon, ``:'', is used to separate constitents in different
sublattices.

After the constituents in the last sublattice a semicolon, ``;''
followed by a single digit, can be used to indicate a degree.  This
degree can have different meanings in different models but is normally
used for the power in a Redlich-Kister series.

A wildcard, ``*'', can be used as constituent in a sublattice
indicating that the parameter is independent of the constituents in
that sublattice.  See also secition \ref{sec:wildcard}.

It is not allowed to have a parameter for an interaction between a
wildcard and a specific element, for example G(FCC,VA,*:VA) which
would represent an interaction $y_{\rm Va}(1-y_{\rm Va}).$

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{The elaborate parameter element}\label{sec:parameter2}

An alternative more elaborated XML element can be used for parameters
which may be preferred by software.  It is straightforward to convert
from one to another.

\bigskip
\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

  Parameter2 & && A more detailed parameter element preferred by software.\\
      & Id & -- & As in TDB file, for example G(LIQUID,A,B:VA;2).  See also the {\bf Parameter} element. \\
      & MPID & H & Model parameter identifier, for example G or TC.  For note see \ref{sec:notes}.\\
      & Phase & -- & Can be omitted inside a phase element, otherwise the phase name.\\
      & LowT & -- & If the default low $T$ limit applies.\\
      & Expr & + & Simple mathematical expression terminated by ;.  If several ranges use a {\bf Trange} element.  See section \ref{sec:expr}.\\
      & HighT & -- & If the default high $T$ limit applies.\\
      & Bibref & + & Bibliographic reference.\\\hline

  ConstArray & && Only inside a {\bf parameter2} element.\\
      & Degree & I & Can be omitted if zero, for note see \ref{sec:notes}.\\\hline
  
  Constituent & && Only inside a {\bf ConstArray} element.\\
      & Sublattice & + &  Sublattice of constituent.\\
      & Species & + & Name of a constituent.\\\hline
\end{tabular}

\bigskip
The {\bf Parameter2} element may be preferred by software but for
manual editing {\bf Parameter} is simpler, see the
appendix~\ref{sec:parameter examples}.  Both elements can be used in
the XTDB file and software can easily convert between them.

If an ``Id'' is present in a {\bf parameter2} element the software
should check that the ``Id'' is identical to the long form and report
an error if not.

The use of wildcards, ``*'', for species in a parameter must be
defined, see section~\ref{sec:wildcard}.


% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{Bibliography for parameters and models.}\label{sec:biblio}

For each parameter there must be a bibliographic reference.  This can
be the same for all parameters from the same source.  The models also
have a reference which should indicate a paper where the model is
explained, see Appendix~\ref{sec:modelex}.

\bigskip
\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

  Bibliography & && Contains bibliographic references\\\hline

  Bibitem & && Only inside a {\bf Bibliography} element.\\
      & Id & + &   Used as bibref attribute for a parameter or model, normally a paper or a comment by the database manager.\\
      & Text & + & Reference to a paper or comment\\
      & Date & -- & Date when the bibref element was created\\
      & Sign & -- & Signature of the one adding the reference\\\hline

\end{tabular}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{XTDB model elements}\label{sec:models}

The XML elements for generally accepted models can be on a separate
file but the model ``Id'' is the important and used in the {\bf
  AmendPhase} element for each phase which has the model.  Most model
has one or more model parameter identifiers (MPID) used in the {\bf
  Parameter} element.  See also appendix~\ref{sec:modelapp}.  Software
specific models should be explained in this XML element.

\bigskip
\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

  Models   & && Contains model descriptions and identifiers used in ``AmendPhase'' elements for phases.  The models usually specify one or more model parameter identifiers needed by the model.\\\hline
\end{tabular}


\subsubsection{XTDB some used model elements}

The XML elements listed here should be defined inside a library of
{\bf Model} elements and their ``Id'' attribute is used to specify the
model parameter identifiers (MPIDs) associated with the model.

\bigskip
\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

%  Models   & && Contains model descriptions and identifiers used in ``AmendPhase'' elements for phases.  The models usually specify one or more model parameter identifiers needed by the model.\\\hline

  Magnetic & && There are several magnetic models.\\
      & Id & + & This Id is used in {\bf AmendPhase} element.\\
      & Refid & + & Where the model is described.\\
      & MPID1 & + & Specifies a magnetic model parameter identifier (MPID) for parameters.\\
      & MPID2 & + & Specifies a magnetic model parameter identifier (MPID) for parameters.\\
      & MPID3 & + & Specifies a magnetic model parameter identifier (MPID) for parameters.\\\hline

  Volume & && Specifies the model for volume of a phase.\\
      & Id & + & This Id is used in the {\bf AmendPhase} element.\\
      & Refid & + & Where the model is described.\\
      & MPID1 & + & Specifies volume model parameter identifier (MPID) for parameters.\\
      & MPID2 & + & Specifies volume model parameter identifier (MPID) for parameters.\\
      & MPID3 & + & Specifies volume model parameter identifier (MPID) for parameters.\\\hline

  FCC4Perm & & & For the FCC 4 sublattice tetrahedron model identical permutations of a parameter are listed only once.\\
     & Id & + & This Id is used in {\bf AmendPhase} element.\\
     & Refid & + & Where the model is described.\\\hline

  BCC4Perm & & & For the BCC 4 sublattice tetrahedron model identical permutations of a parameter are listed only once.\\
     & Id & + & This Id is used in {\bf AmendPhase} element.\\
     & Refid & + & Where the model is described.\\\hline

\end{tabular}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsubsection{New unary models and other models}

There is also a new magnetic model.  Other new {\bf Model} elements
may be included in the XTDB file in order to specify new MPIDs.

\bigskip
\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

  Einstein & && The low $T$ vibrational model.\\
      & Id & + & This Id is used in {\bf AmendPhase} element.\\
      & Refid & + & Where the model is described.\\
      & MPID1 & + & Specifies the Einstein model parameter identifier (MPID) for parameters.\\\hline

  Liquid2state & && The liquid 2-state model.\\
      & Id & + & This Id is used in {\bf AmendPhase} element.\\
      & Refid & + & Where the model is described.\\
      & MPID1 & + & Specifies liquid model parameter identifier (MPID) for parameters.\\
      & MPID2 & + & Specifies Einstein model parameter identifier (MPID) for parameters.\\\hline

  EEC & && Specifies that the Equi-entropy model applies to current database\\
      & Refid & + & Where the model is described.\\\hline

\end{tabular}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsubsection{New XTDB model elements which are constituent specific}\label{sec:toop}

These models are related to specific constituents of a phase and they
must be specified explicitly and cannot be included in the {\bf
  AmendPhase} element for the phase because they include additional
information.

The Toop and Kohler ternary extrapolation models are most likely
specified together with the parameters for a specific ternary.  The
Muggianu ternary model is the default extrapolation model.

There are probably many more models to be considered and we must take
care how they can be integrated in the XML format.

\bigskip
\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

  ToopModel & & & Subset of 3 constituents for which the Toop ternary extrapolation model should be used.    By default the Muggianu extrapolation is used.\\
      & Refid & + & Where the model is described.\\
      & Phase & + & Can be omitted if used inside a {\bf Phase} element.  Otherwise the phase for which the model should be used.\\
      & Constituents & + & Specifies 3 constituents, the Toop constituent first.\\\hline

  KohlerModel & &  & Subset of 3 constituents for which the Kohler ternary extrapolation model should be used.  By default the Muggianu extrapolation is used.\\
      & Refid & + & Where the model is described.\\
      & Phase & + & Can be omitted if used inside a {\bf Phase} element.  Otherwise the phase for which the model should be used.\\
      & Constituents & + & Specifies 3 constituents in any order.\\\hline

\end{tabular}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{Software specific XML elements and others}

The TYPE\_DEFINITION keyword in the TDB file is replaced in this XML
format by {\bf AmendPhase, SplitPhase} and some other attributes.  But
some features of the TYPE\_DEFINITION, which do not relate to the
thermodynamic data, have to be replaced by software specific XML
element.

The ASSESSED\_SYSTEM is an nice feature and I propose we keep such a
possibility also in the XTDB file.

I think it would also be interesting for users to have an XML element
UNASSESSED\_BINARY, especially for encrypted databases, in order to
indicate composition regions where the user should be careful using
calculated results.

The current TYPE\_DEFINITION keyword is sometimes used to create
composition sets if certain elements have been selected or testing if
certain phases should be rejected/restored.  We should think about how
this can be implemented.

\bigskip
\begin{tabular}{|p{0.1\textwidth} p{0.1\textwidth} r p{0.65\textwidth}|}\hline
  Element & Attributes & Note & Explanation\\\hline

  AssessedSystem & & & Indication of system with assessed data.\\
      & Elements & + & The elements or constituents assessed.\\
      & TC & - & Software specific character string to calculate the system with TC.\\
      & Pandat & - & Software specific character string to calculate the system with Pandat.\\
      & PyC & - & Software specific character string to calculate the system with PyCalphad.\\
      & MC & - & Software specific character string to calculate the system with MatCalc.\\
      & OC & - & Software specific character string to calculate the system with OC.\\\hline

  UnassessedBinary & & & Indication of missing binary assessments.\\
       & Elements & + & Two elements with no or incomplete set of assessed
 parameters.\\\hline


 
\end{tabular}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{Notes}\label{sec:notes}

Note that the XML elements and attributes are case sensitive whereas
the names of elements, species, functions and other text are case
insensitive.  For example a chemical element written FE, Fe, fe and fE
is the same element.  Thus CO is Cobalt and C1O is carbonmonoxide.

As a devoted Fortran programmer I prefer character variables with
fixed length and also because it is difficult to format nice output
with very long names of functions or phase names.  For a human it is
also complicated to handle long names, even if they may be
abbreviated.

It is important to be strict here, it is easy to misunderstand a
definition if one is not very explicit.

\begin{description}
\item{\bf A:}\label{sec:noteA} A species name or {\bf Id} must start
  with letter A-Z and can contain letters, digits and the special
  characters ``\_'', ``/'', ``-'' and ``+''.  It must not be
  abbreviated when used as constituent in a {\bf Phase} or {\bf
    Parameter} element.

  Should the characters ``-'' and ``\_'' be treated as identical(?)

\item{\bf B:} The species stoichiometry is a sequence of one or more
  element names (case insensitive) followed by a real number
  specifying the stoichiometric ratio.  Following the TDB standard an
  element with a two letter name does not need a stoichiometric ratio
  equal to unity, for example MGO for MG1O1.  A final stoichiometry
  unity can also be ignored.  No parenthesis are allowed.  For MQMQA
  clusters the stoichiometry can be omitted.

\item{\bf C:} The MQMQA attribute should contain two or more element
  names separated a comma or a sublattice separator ``:'' and
  followed by equal number of unsigned reals representing the bond
  fractions, For example:

  CS,LA:F  9.0 6.0 4.0

  See also appendix~\ref{sec:elementexample}.  According to the MQMQA
  model the cluster stoichometry is calculated as 6.0 divided by the
  bond fractions.  The cluster is always electrically neutral.  For
  endmember clusters (with one species in each sublattice) a real
  representing the SNN/FNN ratio, usually 2.4, must be supplied.

\item{\bf D:} For UNIQUAC species.  Contains two reals representing
  volume and area of the species in m$^3$ and m$^2$ (?).

\item{\bf E:} A {\bf TPfun} name must start with a letter A-Z and may
  contain letters, digits and the special character ``\_''.  It must
  not be longer than 16 characters.  It cannot be abbreviated.

% Add any comments below

%---------------------------------------------------------------
  \newpage
  
\item{\bf F:} A phase name must start with a letter A-Z and have no
  more than 24 characters.  It can contain letters, numbers and the
  special character ``\_''.  In some software``(`` and ``)'' may be
  used in phase names but they can be converted to ``\_'' in other
  software.

  A phase name may be abbreviated in parameters and some other cases
  and thus a phase name must be unique and not an abbreviation of
  another phase.

\item{\bf G:} The configuration model, for example CEF, I2SL, MQMQA
  etc.  Maybe also``RKM'' for a Redlich-Kister-Muggianu model with a
  single lattice and one site?

\item{\bf H:} An initial set of model parameter identifies (MPID) must
  be defined, see section~\ref{sec:mpid} and appendix~\ref{sec:mpid2}.
  At present G is the Gibbs energy, TC Curie $T$, BMAGN the Bohr
  magneton number etc.  The MPID must not be abbreviated.

\item{\bf I:} The degree can be omitted if zero.

\item{\bf K:} The disordered phase within a {\bf SplitPhase} element
  can be omitted if the parameters for the disordered phase are
  provided within the ordered phase.  In OC a suffix ``D'' is added
  after the MPID name and the parameters has the appropriate reduced
  number of sublattices.
  
  For example GD(SIGMA,CR) can be used as {\bf Id} in a {\bf
    Parameter} element for the lattice stability of Cr in SIGMA if the
  SIGMA phase is modelled with 3 or more sublattices and has the {\bf
    SplitPhase} element.  This avoids the need for a meaningless
  disordered DIS\_SIGMA phase.

\item{\bf L:} A BCC or FCC phase with a disordered fraction set (i.e.
  a {\bf SplitPhase} model) can have the ``ordered'' part ``as
  disordered'' subtracted in order to allow the use of the disordered
  part separately.  If so the value of the {\bf Subtract} attribute
  should be ``Y''.

  For phases which never disorder, for example the sigma phase, one
  can also have a ``disordered'' part with a single sublattice
  containing the pure elements as endmembers (and possibly some long
  range interaction parameters).  In Thermo-Calc terminology this is
  the ``NEVER'' model.  In order not to subtract the ordered part as
  disordered in XTDB the {\bf Subtract} attribute can be omitted or
  given any other value than ''Y''.

\end{description}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{The mathematical expression used in Expr}\label{sec:expr}

The mathematical expressions for $T$ and $P$ used in {\bf TPfun,
  Trange} and {\bf Parameter} are the same as in TDB files.  It is
very limited because some software must calculate first and and second
derivatives with respect to $T, P$ (and constitution).  A more
extended mathematical expression could be allowed for expressions
which are not used for database parameters.  In OC it is allowed to
enter more complex expressions for post processing in batch/macro
files.

In the 1990 definition of the TDB file the type of expression allowed
consists of ``simple terms'' such as:

[signed real number] * [{\bf TPfun Id} ] ** [power] *T** [power] *P**[power]

where [power] is an integer (a negative power must be within
parenthesis).  No spaces allowed in a simple term.  A ``complex term''
is a simple term multiplied with a math function of a simple term,
such as:

[simple term] *LN( [simple term] )

An {\bf Expr} attribute in {\bf TPfun, Trange} or {\bf Parameter}
consists one or more terms.  A positive sign of the first term can be
omitted.
  
The following general math functions are allowed in OC:\\ $\exp(),
\ln(), \log(),$ erf().  Note that $\log()$ and $\ln()$ is the same and
erf() is the error function.  The number of math functions can be
extended.

The following math function is needed for the unary
project:\\ GEIN($\theta$) to calculate
$1.5R\theta+3RT\ln(1-\exp(-\theta/T));$

It is not allowed to group several terms together using parenthesis,
for example 5*(1+3*T+2*T*LN(T)).  Either one can multiply all
coefficients with 5 or enter the expression within parentesis as a
{\bf TPfun} and then multiply this with 5.

A square root of $T$ is entered as two {\bf TPfun}://
\begin{verbatim}
<TPfun Id="HALFT" Expr="0.5*LN(T);" />
<TPfun Id="SQRT"  Expr="EXP(HALFT);" />
\end{verbatim}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{The use of wildcards for constituents in parameters}\label{sec:wildcard}

In some parameters a wildcard, ``*'', is used to indicate the
parameter is independent of the constituent in this sublattice.  For
example:

A parameter G(sigma,GA:GA:*;0)=100000; means that the Gibbs energy of
formation of ``phase'' with GA in first and second sublattice is
independent of the constituent in the third.  If one later finds that
for example the Gibbs energy for sigma with TA in the third sublattice
should be just +70000 then, in OpenCalphad, one must set
G(sigma,GA:GA:TA)=-30000; because OC stores the two parameters
separately and will add them when calculating.

This may be handled differently in other software and requires attention.

\subsubsection{The EBEF model use wildcards}

This model use the same notation for parameters as in CEF, and can
replace a large number of endmembers parameters by fewer {\bf
  Parameter} elements, representing bond energies between pairs of
elements.  The bond energy parameters specify constituents in only 2
(or 3) sublattices in phases with 3 or more sublattices and use
wildcards in remaining sublattices.

A parameter G(SIGMA,CR:FE:*:*:*) represent the bond energy between CR
in first and FE in second sublattice, independent on the constituents
in the other sublattices (note that G(SIGMA,FE:CR:*:*:*) is not the
same value!).

Using such parameters (fitted for example to DFT calculated endmember
energies) the number of parameters can be reduced by more than an
order of magnitudes and, in addition, extrapolations can be improved!

\subsubsection{An alternative to wildcards}

To avoid using explicit wildcards one could indicate the sublattice
after the constituent, i.e. to use

G(SIGMA,CR@1:FE@3) instead of G(SIGMA(CR:*:FE:*:*), where the number
after the character ``@'' indicates the sublattice.

\subsubsection{A missing wildcard in the I2SL model parameters}

An I2SL model may exist with only neutrals in the anion sublattice.
For example the elements C and S can be neutrals in the I2SL model and
their parameters are in Thermo-Calc are written as G(I2SL,C), omitting
the cation sublattice.  In OpenCalphad it is written G(I2SL,*:C) which
may be a bit more consistent as it indicates that the constituent on
the cation sublattice is irrelevant.

Alternatively the parameter could be written G(I2SL,C@2) if the
proposal above is accepted.

Note that an interaction between neutrals with a single specific
cation is forbidden in the I2SL model.

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{Model parameter identifiers, MPID}\label{sec:mpid}

An MPID must start with a letter A-Z and contain letters and digits
and not exceed 8? charactes.  It cannot be abbreviated.  There are
already some defined but there should be an extensive list of future
MPIDs to avoid that different software use the same for (slightly)
different things.  See appendix~\ref{sec:mpid2}.
  
The letter ``\&'', frequently used for mobilities, is forbidden in XML
but it can be replaced by some other character, for example ``@''.
Using ``\&amp;'' seems unnecessary and quite clumsy.

\subsection{CVM and the cluster site model}

These should eventually be included also.

% Add any comments below

%---------------------------------------------------------------
\newpage 

\section{Points of discussions}

\begin{enumerate}
\item Should mathematical expressions be terminated by a ``;''?  It is
  not necessary because there is a final double quote of the expression
  but it may be nice.

\item The GEIN function is used in the {\bf Parameter} for a pure
  element modeled with several Einstein $\theta$.  Only one of these
  is selected to vary with composition using the LNTH model parameter
  identifier (MPID).

\item Should the sign of Redlich-Kister terms depend on the
  alphabetrical order of the constituents or as the order the
  constituents are written in the parameter {\bf Id}?  Currently the
  TDB file use alphabetical.

%\item Automatic creation of multiple composition sets when reading the
%  XTDB file?  For example in Fe-Ti-C there will normally be an MC
%  carbide with the FCC model parameters.  Could such a composition set
%  have a different name?  It may be useful to provide handling of
%  carbides or ordering in application software.
%
%  But automatic creation of composition sets will require additional
%  features, for example how to identify the different sets?  Using a
%  ``\#digit'' to identify a composition set?  \#1 is the default
%  (first) composition set.  It also requires a way to specify some
%  limits of the constitution of the composition set, for example att
%  the MC carbide have higher fraction of C (or N) than VA.

\end{enumerate}

\section{Summary}

There are certainly many more things to take care of but I think it is
more important to agree on a minimum common XML format which can make
the thermodynamic databases more useful both for experimentalists,
assessments, database maintenance and thermodynamic software in
particular for the development of new models and applications.  We
have to take one step at a time.

% Add any comments below

%---------------------------------------------------------------
\newpage 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendices
\newpage
\begin{appendices}
\setcounter{equation}{0}
\renewcommand{\theequation}{A\arabic{equation}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

\section{Appendix A: Some examples}\label{sec:examples}

It may not be completely consistent with the definitions above as I
have taken them from my previous conversion software.

\subsection{Elements and species.}\label{sec:elementexample}

The way to define MQMQA constituents and stoichiometry is tentative.

{\small
\begin{verbatim}
  <Element Id="/-" Refstate="ELECTRON_GAS" Mass="  0.000000E+00" />
  <Element Id="VA" Refstate="VACUUM" Mass="  0.000000E+00" />
  <Element Id="AL" Refstate="FCC_A1" Mass="  2.698200E+01" H298="  4.577300E+03" S298="  2.832200E+01" />
  <Element Id="FE" Refstate="BCC_A2" Mass="  5.584700E+01" H298="  4.489000E+03" S298="  2.728000E+01" />
...
  <Species Id="VA" Stoichiometry="VA" />
  <Species Id="AL" Stoichiometry="AL" />
  <Species Id="FE" Stoichiometry="FE" />
  <Species Id="ALFE" Stoichiometry="AL0.5FE0.5" />
  ...
  <Species Id="LA/F" MQMQA="LA:F 6.0 2.0 2.4" />
  <Species Id="LACS/F" MQMQA="LA,CS:F 9.0 6.0 4.0" />
\end{verbatim}
}

The stoichiometry of the MQMQA species is calculated from the bond
ratios according to the model.  The {\bf MQMQA} attribute for an
``endmember'' also include a factor needed in the configurational
entropy expression.

Note that species cannot be abbreviated as constituents in phases or
parameters and thus one can have species which are abbreviations of
another species.  For example ``O'' and ``O2''.

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{Defaults, TPfun and Trange}

Using default $T$ limits the function are not much more complex than
in the TDB file.

{\small
\begin{verbatim}
  <Defaults LowT="298.15" HighT="6000" Refid="U.N.Known" Elements="VA /-" />
...
  <TPfun Id="GHSERAL"  >
    <Trange HighT="700" Expr=" -7976.15+137.093038*T-24.3671976*T*LN(T)-.001884662*T**2-8.77664E-07*T**3+74092*T**(-1);" />
    <Trange HighT="933.47" Expr=" -11276.24+223.048446*T-38.5844296*T*LN(T)+.018531982*T**2  -5.764227E-06*T**3+74092*T**(-1);" />
    <Trange HighT="2900" Expr=" -11278.378+188.684153*T-31.748192*T*LN(T)-1.230524E+28*T**(-9);" />
  </TPfun>
...
  <TPfun Id="LFALFE0"  Expr="-104700+30.65*T;" />
  <TPfun Id="LFALFE1"  Expr="+30000-7*T;" />
  <TPfun Id="LFALFE2"  Expr="+32200-17*T;" />
  <TPfun Id="UFALFE"  Expr="-4000+T;" />
  <TPfun Id="GAL3FE"  Expr="+3*UFALFE+9000;" />
  <TPfun Id="GAL2FE2"  Expr="+4*UFALFE;" />
\end{verbatim}
}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{Phase}\label{sec:phase example}

Entering phases in the XTDB file is a bit more complex but we have get
rid of the TYPE\_DEFINITION.  It is not so nice to use CEF for the
liquid, that is why I think RKM might be a nice option for a liquid
model without sublattices and maybe {\bf Sublattices} can be omitted?

{\small
\begin{verbatim}
 <Phase Id="LIQUID" Configuration="RKM" state="L" >
    <Sublattices Number_of="1"  Ratios="1" >
      <Constituents sublattice="1" list="AL C" />
    </Sublattices>
    <AmendPhase model="LIQ2STATE" />
  </Phase>
...
  <Phase id="A2_BCC" Configuration="CEF" state="S" >
    <Sublattices Number_of="2"  Ratios="1  3" >
      <Constituents sublattice="1" list="AL FE" />
      <Constituents sublattice="2" list="VA" />
    </Sublattices>
    <AmendPhase model="IHJBCC" />
  </Phase>
...
  <Phase id="AL8FE5_D82" Configuration="CEF" state="S" >
    <Sublattices Number_of="2"  Ratios="8  5" >
      <Constituents sublattice="1" list="AL FE" />
      <Constituents sublattice="2" list="AL FE" />
    </Sublattices>
  </Phase>
...
  <Phase id="BCC_4SL" Configuration="CEF" state="S" >
    <Sublattices Number_of="5"  Ratios="0.25  0.25  0.25  0.25  3" >
      <Constituents Sublattice="1" List="AL FE" />
      <Constituents Sublattice="2" List="AL FE" />
      <Constituents Sublattice="3" List="AL FE" />
      <Constituents Sublattice="4" List="AL FE" />
      <Constituents Sublattice="5" List="VA" />
    </Sublattices>
    <SplitPhase Disordered="A2_BCC" Sum="4" Subtract="Y" Refid="09Sun" />
    <AmendPhase model="IHJBCC BCC4Perm" />
  </Phase>

\end{verbatim}
}


For phases one can abbreviate its name and thus one cannot have a
phase with a name which is an abbreviation of another phase name.
Phase names can be abbreviated for each part separated by an
underscore, ``\_''.  A phase name ``AL\_X'' is thus an abbreviation of
``AL2\_X''.

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{Parameters}\label{sec:parameter examples}

{\small
\begin{verbatim}
    <Parameter Id="G(A2_BCC,FE:VA;0)"   Expr="+GHSERFE;" Bibref="91Din" />
    <Parameter Id="TC(A2_BCC,FE:VA;0)"   Expr="1043;" Bibref="91Din" />
    <Parameter Id="BMAGN(A2_BCC,FE:VA;0)"   Expr="2.22;" Bibref="91Din" />
    <Parameter Id="G(AL8FE5_D82,AL:AL;0)"   Expr="+13*GALBCC;" Bibref="08Sun" />
...
    <Parameter Id="G(BCC_4SL,AL:AL:FE:FE:VA;0)"   Expr="+GB2ALFE;" Bibref="08Sun" />
    <Parameter id="G(BCC_4SL,AL,FE:*:*:*:VA;1)"   Expr="-634+0.68*T;" Bibref="08Sun" />
\end{verbatim}
}

In the last parameter above the ``wildcard'' or asterisk, ``*'', is
used for species in three of the sublattices and it means that the
parameter is independent of the constituents in these sublattices.
See the discussion in section~\ref{sec:wildcard} how this is treated.

% Add any comments below

%---------------------------------------------------------------
\newpage 

Using the {\bf Parameter2} element the last two parameters above are:

{\small
\begin{verbatim}
    <Parameter2 Id="G(BCC_4SL,AL:AL:FE:FE:VA;0)" Expr="+GB2ALFE;" Bibref="08Sun" MPID="G" Phase="BCC_4SL" >
       <ConstArray Degree="0">
           <Constituent Sublattice="1" Species="AL" />
           <Constituent Sublattice="2" Species="AL" />
           <Constituent Sublattice="3" Species="FE" />
           <Constituent Sublattice="4" Species="FE" />
           <Constituent Sublattice="5" Species="VA" />
       </ConstArray>
    /Parameter2>
...
    <Parameter2 id="G(BCC_4SL,AL,FE:*:*:*:VA;1)" Expr="-634+0.68*T;" Bibref="08Sun" MPID="G" Phase="BCC_4SL" >
       <ConstArray Degree="1">
           <Constituent Sublattice="1" Species="AL" />
           <Constituent Sublattice="1" Species="FE" />
           <Constituent Sublattice="2" Species="*" />
           <Constituent Sublattice="3" Species="*" />
           <Constituent Sublattice="4" Species="*" />
           <Constituent Sublattice="5" Species="VA" />
       </ConstArray>
    /Parameter2>
\end{verbatim}
}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\subsection{Models}\label{sec:modelex}

An extended form of the {\bf Models} should be defined a separate
global library file with the {\bf Id} and full description but the
models used in an XTDB file should appear in a short form as below.

{\small
\begin{verbatim}
  <Models>
    <Magnetic Id="IHJBCC"  MPID1="TC" MPID2="BMAGN" Bibref="82Her" />
    <Magnetic Id="IHJREST"  MPID1="TC" MPID2="BMAGN" Bibref="82Her" />
    <Einstein Id="GLOWTEIN" MPID1="LNTH" bibref="01Qing" /> 
    <Liquid2state Id="LIQ2STATE" MPID1="G2"  MPID2="LNTH" bibref="14Becker" >
       Unified model for the liquid and the amorphous state treated as an Einstein solid
    </Liquid2state>
    <Volume Id="VOLOWP" MPID1="V0"  MPID2="VA" MPID3="VB" bibref="05Lu" />
  ...
  </Models>
  ...
  <ToopModel Phase="Liquid" Constituents="C Fe Cr" />
\end{verbatim}
}

The {\bf ToopModel} or {\bf KohlerModel} elements can appear anywhere
in the XTDB file, even within the {\bf Phase} element or, more likely
together with all the {\bf Parameter} elements from an assessment of
the ternary C-Fe-Cr system in order to simplify editing a large XTDB
file.  Either way may be a bit complicated to handle by the
calculating software reading the XTDB file.

The {\bf SplitPhase} in example~\ref{sec:phase example} must appear
within a {\bf Phase} element for the ordered phase.

% Add any comments below

%---------------------------------------------------------------
\newpage 

\section{Proposal of Id for models.}\label{sec:modelapp}

The agreed models need only to added by their idenfiers to the phases.

For models that are not agreed a full description of the model and its
model parameter identifiers (MPIDs) should be provided in the XTDB
file.  Note the different software may use different model parameter
identifiers for the same property...

{\tiny
\begin{verbatim}
  <Models>
    <Magnetic id="IHJBCC"  MPID1="TC" MPID2="BMAGN" anti-ferromagnetic_factor=" -1.00" 
       f_below_TC=" +1-0.905299383*TAO**(-1)-0.153008346*TAO**3-.00680037095*TAO**9-.00153008346*TAO**15;"
       f_above_TC=" -.0641731208*TAO**(-5)-.00203724193*TAO**(-15)-.000427820805*TAO**(-25);" bibref="82Her" > in G=f(TAO)*LN(BETA+1) where TAO=T/TC
    </Magnetic>
    <Magnetic id="IHJREST"  MPID1="TC" MPID2="BMAGN" anti-ferromagnetic_factor=" -3.00" 
       f_below_TC=" +1-0.860338755*TAO**(-1)-0.17449124*TAO**3-.00775516624*TAO**9-.0017449124*TAO**15;"
       f_above_TC=" -.0426902268*TAO**(-5)-.0013552453*TAO**(-15)-.000284601512*TAO**(-25);" bibref="82Her" > in G=f(TAO)*LN(BETA+1) where TAO=T/TC
    </Magnetic>
    <Magnetic id="IHJQX"  MPID1="CT" MPID2="NT" MPID3="BMAGN" anti-ferromagnetic_factor="  0.00" 
       f_below_TC=" +1-0.842849633*TAO**(-1)-0.174242226*TAO**3-.00774409892*TAO**9-.00174242226*TAO**15-.000646538871*TAO**21;"
       f_above_TC=" -.0261039233*TAO**(-7)-.000870130777*TAO**(-21)-.000184262988*TAO**(-35)-6.65916411E-05*TAO**(-49);" bibref="12Xiong" > in G=f(TAO)*LN(BETA+1) where TAO=T/CT or T/NT
    </Magnetic>
    <Einstein id="GLOWTEIN" MPID1="LNTH" bibref="01Qing" > 
       Gibbs energy due to the Einstein low T vibrational entropy model, G=1.5*R*THETA+3*R*T*LN(1-EXP(-THETA/T)).
    </Einstein>
    <Liquid2state id="LIQ2STATE" MPID1="G2"  MPID2="LNTH" bibref="14Becker" >
       Unified model for the liquid and the amorphous state treated as an Einstein solid
    </Liquid2state>
    <Volume id="VOLOWP" MPID1="V0"  MPID2="VA" MPID3="VB" bibref="05Lu" >
       The volume of a phase is described as function of T, P and its constitution.
    </Volume>
    <SplitPhase Disordered=" " Sum=" " Subtract=" " Refid="09Sun" >
       The disordered fractions are summed over the ordered sublattices indicated by Sum.  The Gibbs energy calculated 2 or 3 times as indicated by the CEF_appendix but the Subtract but the configurational entropy only once.
    </SplitPhase>
    <FCC-permutations id="FCC4Perm" bibref="09Sun" >
       Permutations of ordered FCC parameters with the same set of elements are listed only once.
    </FCC-permutations>
    <BCC-permutations id="BCC4Perm" bibref="09Sun" >
       Permutations of ordered BCC parameters with the same set of elements are listed only once.
    </BCC-permutations>
    <EEC id="EEC" bibref="20Sun" >
       Equi-Entropy Criterion means that solid phases with higher entropy that the liquid phase must not be stable.
    </EEC>
  </Models>
\end{verbatim}
}

% Add any comments below

%---------------------------------------------------------------
\newpage 

\section{Proposal of initial set of Model Parameter Identifiers, MPID}\label{sec:mpid2}

The 32 model parameter identifiers, MPID, defined in OC as shown in
Table~\ref{tab:mpis}.  In OC a parameter for a disordred part of a
{\bf SplitPhase} use the same phase name but the MPID has a suffix
``D''.

\begin{table}[!h]
  \caption{Current set of model parameter identifiers in OC.  For each
    parameter it is indicated if it can depend on $T$, $P$ or have an
    extra constituent specification.  Most of them have no associated
    code.}\label{tab:mpis}

{\small
\begin{verbatim}
Indx Ident T P Specification                Status Note
   1 G     T P                                   0 Gibbs Energy
   2 TC    - P                                   2 Combined Curie/Neel T
   3 BMAG  - -                                   1 Average Bohr magneton number
   4 CTA   - P                                   2 Curie temperature
   5 NTA   - P                                   2 Neel temperature
   6 IBM   - P &<constituent#sublattice>;       12 Individual Bohr magneton number
   7 LNTH  - P                                   2 Einstein temperature
   8 V0    - -                                   1 Volume at T0, P0
   9 VA    T -                                   4 Thermal expansion
  10 VB    T P                                   0 Bulk modulus
  11 VC    T P                                   0 Alternative volume parameter
  12 VS    T P                                   0 Diffusion volume parameter
  13 MQ    T P &<constituent#sublattice>;       10 Mobility activation energy
  14 MF    T P &<constituent#sublattice>;       10 RT*ln(mobility freq.fact.)
  15 MG    T P &<constituent#sublattice>;       10 Magnetic mobility factor
  16 G2    T P                                   0 Liquid two state parameter
  17 THT2  - P                                   2 Smooth step function T
  18 DCP2  - P                                   2 Smooth step function value
  19 LPX   T P                                   0 Lattice param X axis
  20 LPY   T P                                   0 Lattice param Y axis
  21 LPZ   T P                                   0 Lattice param Z axis
  22 LPTH  T P                                   0 Lattice angle TH
  23 EC11  T P                                   0 Elastic const C11
  24 EC12  T P                                   0 Elastic const C12
  25 EC44  T P                                   0 Elastic const C44
  26 UQT   T P &<constituent#sublattice>;       10 UNIQUAC residual parameter (OC)
  27 RHO   T P                                   0 Electric resistivity
  28 VISC  T P                                   0 Viscosity
  29 LAMB  T P                                   0 Thermal conductivity
  30 HMVA  T P                                   0 Enthalpy of vacancy formation (MatCalc)
  31 TSCH  - P                                   2 Schottky anomaly T (OC)
  32 CSCH  - P                                   2 Schottky anomaly Cp/R. (OC)
  33 NONE  T P                                   0 Unused
\end{verbatim}
  }
\end{table}
 
% Add any comments below

%---------------------------------------------------------------
\newpage

\section{Replies}

If you have added notes on any pages please indicate who you are and
make a summary here.

\bigskip
{\Large
\begin{tabular}{|l|l||l|l||l|l||l|l|}\hline
Name      & Reply & Name   & Reply & Name  & Reply & Name  & Reply \\\hline 
  Abe     &      & Alex    &      & Bengt   &      &  Bosse   &  \\\hline
  Erwin   &      & Fabio   &      & Fan     &      &  Lina    &  \\\hline
  Malin   &      & Nath    &      & Reza    &      &  Richard &   \\\hline
  Shuanglin &    & Ursula  &      &         &      &          &   \\\hline
\end{tabular}
}

\end{appendices}
\end{document}

\newpage
\setcounter{equation}{0}
\renewcommand{\theequation}{B\arabic{equation}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{B\arabic{figure}}

\section{Replies about the proposal}

Summary of comments.

\bigskip
\begin{tabular}{|l|l|}
 Contributer  & Comments \hspace{120mm} \\\hline
  Abe     & \\
          & \\
          & \\\hline
  Alex    & \\
          & \\
          & \\\hline
  Bengt   & \\
          & \\
          & \\\hline
  Bosse   & \\
          & \\
          & \\\hline
  Erwin   & \\
          & \\
          & \\\hline
  Fabio   & \\
          & \\
          & \\\hline
  Fan     & \\
          & \\
          & \\\hline
  Lina    & \\
          & \\
          & \\\hline
  Malin   & \\
          & \\
          & \\\hline
  Nath    & \\
          & \\
          & \\\hline
  Reza    & \\
          & \\
          & \\\hline
  Richard & \\
          & \\
          & \\\hline
  Shuanglin & \\
          & \\
          & \\\hline
  Ursula  & \\
          & \\
          & \\\hline
  Anyone else & \\
          & \\
          & \\\hline
\end{tabular}
\end{appendices}

\end{document}


\bigskip
{\Large
\begin{tabular}{|l|l||l|l||l|l||l|l|}\hline
Name      & Reply & Name   & Reply & Name  & Reply & Name  & Reply \\\hline 
  Abe     &      & Alex    &      & Bengt   &      &  Bosse   &  \\\hline
  Erwin   &      & Fabio   &      & Fan     &      &  Lina    &  \\\hline
  Malin   &      & Nath    &      & Reza    &      &  Richard &   \\\hline
  Shuanglin &    & Ursula  &      &         &      &          &   \\\hline
\end{tabular}
}

\newpage

